---
title: "Using ploidyverseClasses in your R package"
author: "Lindsay V. Clark, University of Illinois, Urbana-Champaign"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Using ploidyverseClasses in your R package}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Purpose of this document

This document is intended for R package developers who wish to utilize S4
classes from the ploidyverse in order to make their package interoperable
with other ploidyverse packages.  It assumes basic familiarity with the process
of creating an R package.

## How to indicate dependency on `ploidyverseClasses`

The first question you should ask yourself is, will the `ploidyverseVCF`
class be mandatory for using your package?  Or, will users frequently
want to import or export data in a different format, and not care about
incorporating other ploidyverse software into their workflow?

`ploidyverseClasses` depends on the Bioconductor package 
[`VariantAnnotation`](https://bioconductor.org/packages/release/bioc/html/VariantAnnotation.html).
Like many Bioconductor packages, `VariantAnnotation` can be unavoidably 
cumbersome to install.  We on the ploidyverse core team believe that it is
worthwhile for its flexibility in importing, storing, and exporting genomic
variants and metadata, but we want to minimize how much we force it onto people
who don't want it.

### Option A: mandatory installation of `ploidyverseClasses`

In this case, you will indicate dependency on `ploidyverseClasses` in the
standard way that you would for any R package.  In your `DESCRIPTION` file,
list `ploidyverseClasses` and `methods` on the `Imports` line.

```
Imports: ploidyverseClasses, methods
```

In the  `NAMESPACE` file, you should include the lines:

* `importClassesFrom("ploidyverseClasses", "ploidyverseVCF")`
* an `importFrom` line for any functions you are using from `ploidyverseClasses`
* `importFrom` lines for any functions you are using from `VariantAnnotation` 
or other packages

For any ploidyverse or Bioconductor functions that you use, I recommend looking
at the help page to determine exactly which package it originated from.  

Remember that any package with an `importFrom` line will also need a mention 
in the `Imports` line of your `DESCRIPTION` file.

### Option B: optional installation of `ploidyverseClasses`

If you don't want to force someone to install `ploidyverseClasses` (and its
dependencies) in order to use your package, you should instead list
`ploidyverseClasses` and `methods` on the `Suggests` line of your `DESCRIPTION`
file.  Also list any Bioconductor packages whose functions you want to use
in `Suggests`.

```
Suggests: ploidyverseClasses, methods
```

Then within your function definitions, use the `::` operator to indicate
the package that a function comes from.  For example, to use the `seqinfo`
function, type `GenomeInfoDb::seqinfo`.

```{r eval = FALSE}
# a trivial example
GetSeqInfo <- function(x){
  GenomeInfoDb::seqinfo(x)
}
```


Note to self: Need to make sure S4 classes can import this way.